### 2025-08-30 face_recognition/docs/6_conclucion.tex

\section{Conclusión}

% reunirnos mañana para elaborarla juntos o tambien internamente puede avanzar con los demas, y pulir detalles.