### 2025-08-31 face_recognition/docs/6_conclucion.tex

\section{Conclusión}

TODO: we need to update this section to summarize the deployment and pipeline sections.